Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Chad Bown, Nuno Limao, Giovanni Maggi, Anson Soderbery, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at SITE for very helpful comments and discussions.}

### 2017-01-16 draft.tex
finish introducing and justifying -e

The per-period welfare function of the home government is given by
\begin{equation}
  W(\ga(s,e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(s,e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau) - e
  \label{eq:wel}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\pi$ represents profits, $\ga(s,e)$ is the weight placed on profits (producer surplus) in the import-competing industry, $s$ is an exogenous state variable, $e$ is lobbying effort in terms of the numeraire, and $\mathit{TR}$ is tariff revenue. Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, may be affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{While the standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them. See Section~\ref{sec:objfcn} for a detailed discussion.} I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to the case where political pressure  results from the rent-seeking behavior of lobbies only.

>>>>

The per-period welfare function of the home government is given by
\begin{equation}
  W(\ga(s,e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(s,e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau) - e
  \label{eq:wel}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\pi$ represents profits, $\ga(s,e)$ is the weight placed on profits (producer surplus) in the import-competing industry, $s$ is an exogenous state variable, $e$ is lobbying effort in terms of the numeraire, and $\mathit{TR}$ is tariff revenue. Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, may be affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{While the standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them. See Section~\ref{sec:objfcn} for a detailed discussion.} I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to the case where political pressure  results from the rent-seeking behavior of lobbies only. I subtract the cost of lobbying effort from the government's objective function because if stands to reason that if the government cares about lobbying gross profits, it should also care about the costs.\footnote{Aside from making the government's optimum interior in $\ga$, taking account of the cost of lobbying changes the results that follow quantitatively but not qualitatively because there is no marginal impact on the government's decision-making regarding the optimal applied tariff.} This is often abstracted from in the closely-related literature, but it seems important to take into account in this environment of non-transferable utility and decreasing returns to lobbying. Note that there is no cost of lobbying effort in models that assume political pressure is exogenous only.
